---
title: "sds-hw9"
output: html_document
date: "2025-04-19"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, # show the code
                      eval = TRUE, # run the code
                      warning = TRUE, # show warning messages
                      message = FALSE, # show basic messages
                      fig.align = "center"
                      )
```

```{r}
# Problem 1)

# 1. Load libraries and data
library(ggplot2)
library(broom)
library(dplyr)
library(readr)
library(tidyverse)

solder <- read.csv("solder.csv")
groceries <- read_csv('groceries.csv')

# 2. Convert to factors (don’t hardcode levels)
solder$Opening <- factor(solder$Opening)
solder$Solder  <- factor(solder$Solder)

# 3. Diagnostic: make sure each factor has ≥2 levels
print(table(solder$Opening))  # should show at least two non‑zero counts
print(table(solder$Solder))

# 4. Part A: Exploratory plots

# 4a. Skips vs. Opening
ggplot(solder, aes(Opening, skips)) +
  geom_boxplot(fill = "lightblue") +
  geom_jitter(width = 0.1, alpha = 0.5) +
  labs(
    title = "Number of Solder Skips vs. Gun Opening Size",
    x     = "Opening Size",
    y     = "Number of Skips"
  ) +
  theme_minimal()
# Caption: Boards with medium-sized openings show fewer skips on average than small or large.

ggplot(solder, aes(Solder, skips)) +
  geom_boxplot(fill = "salmon") +
  geom_jitter(width = 0.1, alpha = 0.5) +
  labs(
    title = "Number of Solder Skips vs. Alloy Thickness",
    x     = "Alloy Thickness",
    y     = "Number of Skips"
  ) +
  theme_minimal()
# Caption: Thin alloy tends to produce fewer skips compared to thick alloy.
```

```{r}

# Part B: Fit regression with interaction
mod1 <- lm(skips ~ Opening * Solder, data = solder)


# Part B: Table of estimates & 95% CIs
tbl1 <- tidy(mod1, conf.int = TRUE)
print(tbl1[, c("term", "estimate", "conf.low", "conf.high")])

```

# Part C: Interpret coefficients 
# With 95% confidence, The true baseline number of solder skips on AT&T's
#circuit boards when the soldering gun opening is large and the soldering
#alloy is thick is between -0.63 and 1.41.

#With 95% confidence, The true main effect of the soldering gun's opening
#being medium-sized lies between 0.96 and 3.85 additional solder skips.

#With 95% confidence, The true main effect of the soldering gun's opening
#being small-sized lies between 3.68 and 6.57 additional solder skips.

#With 95% confidence, The true main effect of the soldering alloy being
#thing lies between 0.84 and 3.72 additional solder skips.

#With 95% confidence, The true interaction effect of the soldering gun
#opening being medium-sized and the soldering alloy being thin lies
#between -2.78 and 1.30 additional solder skips.

#With 95% confidence, The true interaction effect of the soldering gun
#opening being small-sized and the soldering alloy being thin lies
#between 7.61 and 11.70 additional solder skips.


```{r}
# Part D: Predicted means for each factor combination
newdata <- expand.grid(
  Opening = levels(solder$Opening),
  Solder  = levels(solder$Solder)
)
newdata$pred_skips <- predict(mod1, newdata)
print(newdata)
```

```{r}
#  Problem 2)
groc <- read_csv("groceries.csv")

groceries %>% 
  group_by(Store) %>% 
  summarize(mean_price = mean(Price)) %>%
  # Ordering stores by price
  mutate(Store = fct_reorder(Store, mean_price)) %>% 
  ggplot(aes(x = Store, y = mean_price)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  coord_flip() +
  labs(x = "Store",
       y = "Products' Mean Price",
       title = "Mean Product Price by Texas Grocery Stores")

```

```{r}
# Part B: number of stores per product
count_by_product <- groc %>%
  distinct(Store, Product) %>%
  count(Product, name = "n_stores") %>%
  arrange(n_stores)

ggplot(count_by_product, aes(x = n_stores, y = fct_reorder(Product, n_stores))) +
  geom_col(fill = "coral") +
  labs(
    title = "Number of Stores Selling Each Product",
    x     = "Number of Stores",
    y     = "Product"
  ) +
  theme_minimal()
```

```{r}
# Part C: Price ~ Product + Type
tbl_type <- tidy(mod_type, conf.int = TRUE)

# List out all the store‐type terms:
tbl_type %>%
  filter(str_detect(term, "^Type")) %>%
  select(term, estimate, conf.low, conf.high)
```

```{r}
# Part D: Price ~ Product + Store
mod_store <- lm(Price ~ Product + Store, data = groc)
tbl_store <- tidy(mod_store, conf.int = TRUE)
# two lowest & two highest store effects:
tbl_store %>%
  filter(str_detect(term, "^Store")) %>%
  arrange(estimate) %>%
  slice(1:2)  # lowest
tbl_store %>%
  filter(str_detect(term, "^Store")) %>%
  arrange(desc(estimate)) %>%
  slice(1:2)  # highest
```

```{r}
### **Part E: Price Discrimination**

groceries$Store <- relevel(factor(groceries$Store), ref = "H-E-B")

# refit the model
store_lm1 <- lm(Price ~ Product + Store, data = groceries)
tidy_store <- tidy(store_lm1)

coef(store_lm)["StoreCentral Market"]

store_terms <- tidy_store[grep("^Store", tidy_store$term), ]

#cheapest store vs HEB
store_terms <- store_terms[order(store_terms$estimate), ]

cheapest_store <- store_terms[1, ]

print(cheapest_store)

#priciest store vs HEB
store_terms <- store_terms[order(store_terms$estimate), ]

priciest_store <- tail(store_terms,1)

print(priciest_store)
```

```{r}
# Part F: Price ~ Product + Income10K
groc2 <- groc %>% mutate(Income10K = Income / 10000)
mod_inc <- lm(Price ~ Product + Income10K, data = groc2)
tbl_inc <- tidy(mod_inc)
est_inc  <- tbl_inc %>% filter(term == "Income10K") %>% pull(estimate)
sd_inc   <- sd(groc2$Income10K, na.rm = TRUE)
sd_price <- sd(groc2$Price,     na.rm = TRUE)
std_beta <- est_inc * (sd_inc / sd_price)
cat("Income10K coef:", round(est_inc,3), 
    "\n1‑SD income →", round(std_beta,3), "SD price\n")
```
#Problem 3) 
1.True. Figure A1 shows a clear upward trend, and the simple regression of policies ~ minority gives a slope of 0.014 (95% CI [0.009, 0.018]), with R² ≈ 0.52, indicating a moderate positive association

2. Undecidable. No model includes a minority x age interaction. The only regression relating those two is minority ~ age (R²=0.061), which treats them as separate predictors without testing their interplay.

3. False. In the interaction model (policies ~ minority * fire_risk), the slope in high‑risk areas is 0.010, and the low‑risk slope is 0.010 + (–0.001) = 0.009. The interaction term (–0.001) is essentially zero and not significant (95% CI [–0.012, 0.010], p = 0.839), so there’s no evidence of a stronger effect in high‑risk ZIP codes ​

4. False. The unadjusted model (policies ~ minority) gives a coefficient of 0.014 (95% CI[0.009, 0.018]). Adding income reduces it only to 0.010 (95% CI [0.004, 0.015]), but it remains positive and significant. Income does not fully “explain away” the minority effect

5. True. In the full model (undefined), the minority coefficient remains positive at 0.008 (95 % CI [0.003, 0.014], p = 0.006), showing the association persists after adjusting for all controls 
